Adaptive Video Streaming with Optimized Bitstream Extraction and PID-based Quality Control

### 2015-12-04 tmm.tex
Move description of how the system works to last sub-section of experiments

To implement our proposed adaptive streaming system requires the combination of the optimized bitstream extraction and the PID-based quality control algorithm. The latter determines the current bitrate to follow the bandwidth change, and the former will ensure the sub-stream with the best possible quality is obtained and sent under that bitrate constraint. They work together to improve the user's experience of watching videos online. With DSS as the implementation basis, we have integrated the proposed algorithms into it for the purpose of both research and application. The system works as follows. The greedy-like algorithm present in \ref{subsec:priority-assign} is performed first to determine priority values for all enhancement NALUs in an SVC bitstream. Then the priority values are injected into the bitstream and stored in the priority identifier field of the NALU header. When DSS streams this video, it can extract an optimized sub-stream by simply checking this field and discarding the packets with lower priority values. The bitrate of the sub-stream to be extracted is determined by the PID-based quality control algorithm according to network condition.

@@

The performance comparisons for City, Foreman, Harbour, and Soccer are shown in Fig. \ref{fig:extraction-performance}. It can be seen that the R-D curves of the proposed method are always on the top (with only one exception in Foreman). For each sequence, the maximum and average PSNR gain through all 10 bitrate points are calculated and listed in Table \ref{tab:extraction-gain}, which shows that our extraction method can achieve a significant performance gain compared with JSVM Basic, and also performs better than JSVM QL. Table \ref{tab:extraction-gain} also contains results when the optimization window size is set to 48 frames, i.e., 6 GOPs. It is demonstrated that a larger window size can increase performance to some degree. However, it should be noted that, in practice the optimization window may not be that large, for the reason stated at the end of sub-section \ref{subsec:priority-assign}.

@@

After combining the two algorithms together, the proposed adaptive streaming system is able to deliver a higher and smoother video quality for the user. We compare the combined system's performance with DSS that supports SVC but has its original quality control method and JSVM Basic bitstream extraction. The result for the test sequence Soccer is present in Fig. \ref{fig:fluctuation-psnr}. It shows the streamed video's PSNR change when available bandwidth fluctuates around certain values (400kbps, 600kbps, 800kbps and 1000kbps). It can be seen that, the proposed system performs better in that it has effectively avoided unnecessary quality drop (e.g., note the curve around time 350s in Fig. \ref{fig:Fluctuation-PSNR3} or 150s in Fig. \ref{fig:Fluctuation-PSNR4}) and been able to keep a relatively higher PSNR value (e.g., note the curve around time 450s in Fig. \ref{fig:Fluctuation-PSNR1} or 500s in Fig. \ref{fig:Fluctuation-PSNR2}). The average and variance value of PSNR through the time is also calculated and present in Fig. \ref{fig:fluctuation-psnr}, showing clearly the proposed system's advantage. For instance, in Fig. \ref{fig:Fluctuation-PSNR3} where bandwidth fluctuates around 800kbps, the average PSNR delivered by the proposed system is 38.75dB, 0.83dB higher than that of the baseline system, and the PSNR variance is also reduced from 1.24 to 0.69, indicating a more smooth video quality change.

>>>>

To implement our proposed adaptive streaming system requires the combination of the optimized bitstream extraction and the PID-based quality control algorithm. The latter determines the current bitrate to follow the bandwidth change, and the former will ensure the sub-stream with the best possible quality is obtained and sent under that bitrate constraint. They work together to improve the user's experience of watching videos online. With DSS as the implementation basis, we have integrated the proposed algorithms into it for the purpose of both research and application.

@@

The performance comparisons for City, Foreman, Harbour, and Soccer are shown in Fig. \ref{fig:extraction-performance}. It can be seen that the R-D curves of the proposed method are always on the top (with only one exception in Foreman). For each sequence, the maximum and average PSNR gain through all 10 bitrate points are calculated and listed in Table \ref{tab:extraction-gain}, which shows that our extraction method can achieve a significant performance gain compared with JSVM Basic, and also performs better than JSVM QL. Table \ref{tab:extraction-gain} also contains results when the optimization window size is set to 48 frames, i.e., 6 GOPs. It is demonstrated that a larger window size can increase performance to some degree. However, it should be noted that, in practice the optimization window may not be that large, for the reason stated at the end of Section \ref{subsec:priority-assign}.

@@

After combining the two algorithms together, the proposed adaptive streaming system is able to deliver a higher and smoother video quality for the user. The system works as follows. The greedy-like algorithm present in \ref{subsec:priority-assign} is performed first to determine priority values for all enhancement NALUs in an SVC bitstream. Then the priority values are injected into the bitstream and stored in the priority identifier field of the NALU header. When DSS streams this video, it can extract an optimized sub-stream by simply checking this field and discarding the packets with lower priority values. The bitrate of the sub-stream to be extracted is determined by the PID-based quality control algorithm according to the network condition.

We compare the combined system's performance with DSS that supports SVC but has its original quality control method and JSVM Basic bitstream extraction. The result for the test sequence Soccer is present in Fig. \ref{fig:fluctuation-psnr}. It shows the streamed video's PSNR change when available bandwidth fluctuates around 400kbps, 600kbps, 800kbps and 1000kbps (bandwidth fluctuation is simulated the same way as in the previous experiments). It can be seen that, the proposed system performs better in that it has effectively avoided unnecessary quality drop (e.g., note the curve around time 350s in Fig. \ref{fig:Fluctuation-PSNR3} or 150s in Fig. \ref{fig:Fluctuation-PSNR4}) and been able to keep a relatively higher PSNR value (e.g., note the curve around time 450s in Fig. \ref{fig:Fluctuation-PSNR1} or 500s in Fig. \ref{fig:Fluctuation-PSNR2}). The average and variance value of PSNR through the time is also calculated and present in Fig. \ref{fig:fluctuation-psnr}, showing clearly the proposed system's advantage. For instance, in Fig. \ref{fig:Fluctuation-PSNR3} where bandwidth fluctuates around 800kbps, the average PSNR delivered by the proposed system is 38.75dB, 0.83dB higher than that of the baseline system, and the PSNR variance is also reduced from 1.24 to 0.69, indicating a more smooth video quality change.
